---
title: "simpleRegress"
author: "Matias Lucha"
date: "Summer 2024"
output: html_document
---

dont do length dont do month

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

For model length~neck, adjusted r^2 = 0.75

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age
```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

adjusted r^2 for the age model is 0.51, this is less than the adjusted r^2 for neck, so this model is not as good as the neck model.

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

adjusted r^2 for the age model is 0.79, this is less than the adjusted r^2 for neck, so this model is not as good as the neck model.

```{r}
headwidthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
summary.lm(headwidthModel)
```

Headwith r^2 is 0.5321 which is not as well as other values. 

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```

Sex r^2 value is 0.08373 which is the lowest score of all.


```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```

Headlength r^2 value is is 0.843 which is the best value. 

```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

Weight r^2 value is 0.743 which is the 2nd best value. 

In conclusion, the model LENGTH~HEADLEN is the best indicator comparison with r^2 is 0.843. Followed by 0.7869 with CHEST, 0.743 with WEiGHT, 0.5321 in HEADWIDTH, 0.5073 in AGE, and lastly SEX with a value of 0.084.